Supporting Information:\\Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-16 Manuscript/SI.tex

\begin{figure}[!hp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
  \caption{\label{fig:ordPars_waterModels}
    Changes of head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of POPC bilayer
    as a function of CaCl$_2$ concentrations in bulk ($C_b$) from ECC-POPC simulations at 313K with different
    water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
    together with experimental data (DPPC (323~K) \cite{akutsu81} and POPC (313~K) \cite{altenbach84}). 
    Ion concentrations ($C_b$) are calculated from the cation number density $C_{np}$
    at the farthest point from the lipid bilayer in the aqueous phase as [ion]=$C_{np}/0.602$.
  }
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl_si}
    Changes of head group and glycerol carbon $g_3$ order parameters, and the P-N vector orientation of POPC bilayer
    as a function of NaCl concentrations in bulk ($C_b$)
    from simulations with different force fields at 313~K together with experimental
    data (DPPC (323~K)~\cite{akutsu81} and POPC (313~K)~\cite{altenbach84}). 
%    Ion concentrations in bulk water are shown in x-axis. 
%    Values from simulations are calculated from the of cation number density $C_{np}$
%    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters at 293~K is taken directly
    from Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
  }
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_ECC-lipids_346mM_CaCl.pdf} \\
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_Charmm36_450mM_CaCl_Matti.pdf}
  \caption{\label{fig:hist_residence_times}
   Histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer
   from ECC-POPC (top) and CHARMM36 (bottom) simulations with ECC-ions.
   Both simulations had the same concentration of \ce{CaCl_2} respect to water ($C_w$= 450~mM).
   CHARMM36 simulation was directly taken from Refs.~\citenum{javanainen17,zenodo.259376}.
   Scales of x-axes represent the lengths of the simulations used for analysis.
   In ECC-POPC simulation, 90\% of the residence times are
   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
   with the longest observed residence time being $141\,\mathrm{ns}$,
   which is well below the total length of the simulation (200~ns).
   This is, however not the case in CHARMM36 simulation,
   where residence times of several calcium cations are    
   apparently limited by the length of the simulation.
   %With such a simulation, which is relatively short for the employed model,
   %we can merely estimate an upper bound that
   %only
   Less than 60\% of the residence times are
   %is formed by cations interacting with the phospholipid membrane
   shorter than the half of the simulation length~(400~ns)
   in CHARMM36 simulation.
  }
\end{figure}

@@

\begin{figure}[!p]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_cacl_GMX_oMM_compar}
    Changes of the head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of a POPC bilayer 
    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
    from Lipid14 \cite{dickson14} and ECC-POPC simulations ran
    with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} at 313~K
    together with experiments (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
    %Ion concentrations in bulk water are shown in x-axis. 
    Bulk concentrations from simulations are calculated 
    from the farthest point from the lipid bilayer in the aqueous phase
    with an error estimate of 10\,mM.
  }
\end{figure}

>>>>

\begin{figure}[!hp]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
  \caption{\label{fig:ordPars_waterModels}
    Changes of head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of POPC bilayer
    as a function of CaCl$_2$ concentrations in bulk ($C_{ion}$) from ECC-POPC simulations at 313K with different
    water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
    together with experimental data (DPPC (323~K) \cite{akutsu81} and POPC (313~K) \cite{altenbach84}). 
    Ion concentrations ($C_{ion}$) are calculated from the cation number density $C_{np}$
    at the farthest point from the lipid bilayer in the aqueous phase as [ion]=$C_{np}/0.602$.
  }
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl_si}
    Changes of head group and glycerol carbon $g_3$ order parameters, and the P-N vector orientation of POPC bilayer
    as a function of NaCl concentrations in bulk ($C_{ion}$)
    from simulations with different force fields at 313~K together with experimental
    data (DPPC (323~K)~\cite{akutsu81} and POPC (313~K)~\cite{altenbach84}). 
%    Ion concentrations in bulk water are shown in x-axis. 
%    Values from simulations are calculated from the of cation number density $C_{np}$
%    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters at 293~K is taken directly
    from Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
  }
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_ECC-lipids_346mM_CaCl.pdf} \\
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_Charmm36_450mM_CaCl_Matti.pdf}
  \caption{\label{fig:hist_residence_times}
   Histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer
   from ECC-POPC (top) and CHARMM36 (bottom) simulations with ECC-ions.
   Both simulations had the same concentration of \ce{CaCl_2} respect to water ($C_{ion}'$= 450~mM).
   CHARMM36 simulation was directly taken from Refs.~\citenum{javanainen17,zenodo.259376}.
   Scales of x-axes represent the lengths of the simulations used for analysis.
   In ECC-POPC simulation, 90\% of the residence times are
   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
   with the longest observed residence time being $141\,\mathrm{ns}$,
   which is well below the total length of the simulation (200~ns).
   This is, however not the case in CHARMM36 simulation,
   where residence times of several calcium cations are    
   apparently limited by the length of the simulation.
   %With such a simulation, which is relatively short for the employed model,
   %we can merely estimate an upper bound that
   %only
   Less than 60\% of the residence times are
   %is formed by cations interacting with the phospholipid membrane
   shorter than the half of the simulation length~(400~ns)
   in CHARMM36 simulation.
  }
\end{figure}

@@

\begin{figure}[!p]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_cacl_GMX_oMM_compar}
    Changes of the head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of a POPC bilayer 
    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
    from Lipid14 \cite{dickson14} and ECC-POPC simulations ran
    with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} at 313~K
    together with experiments (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
    %Ion concentrations in bulk water are shown in x-axis. 
    Bulk concentrations from simulations are calculated 
    from the farthest point from the lipid bilayer in the aqueous phase
    with an error estimate of 10\,mM.
  }
\end{figure}
